---
title: "Case Study 2"
author: "Marc Brooks (Presenter), Bo Liu (Checker), Shirley Mathur (Programmer), Aasha Reddy (Writer)"
date: "11/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      fig.align = 'center')
```

```{r, warning=FALSE}
library(tidyverse)
library(lme4)
library(rstan)
library(brms)
library(knitr)
library(kableExtra)
library(patchwork)
library(lubridate)
library(gridExtra)
library(influence.ME)
library(pROC)
library(tidybayes)


options(scipen = 0, digits = 4)
ggplot2::theme_set(ggplot2::theme_bw())
```


```{r}
#load data
registered <- read.delim("../data/voter_stats_20201103.txt")
voted <- read.delim("../data/history_stats_20201103.txt")
```

# Introduction 

The North Carolina State Board of Elections (NCSBE) is the agency charged with the administration of the elections process and campaign finance disclosure and compliance. Among other things, they provide voter registration and turnout data online. Our goal in this case study is to use the NC voter files for the general elections in November 2020 to identify/estimate how different groups voted in the 2020 elections (out of those registered).

## Research Questions 

1) How did different demographic subgroups vote in the 2020 general elections? For example, how did the turnout for males compare to the turnout for females after controlling for other potential predictors?

2) Did the overall probability or odds of voting differ by county in 2020? Which counties differ the most from other counties?

3) How did the turnout rates differ between females and males for the different party affiliations?

4) How did the turnout rates differ between age groups for the different party affiliations?

# Data and cleaning 
We have two datasets that we will merge: voter_stats_20201103.txt contains information about the aggregate counts of registered voters by the demographic variables, and history_stats_20201103.txt contains information about the aggregate counts of voters who actually voted by the demographic variables. We first clean the data. 

Our outcome variable of interest needs to be created. We are examining the turnout of voters, and we have counts of total registered voters in one dataset (we will call this dataset "registered"), and counts of people who actually voted in another dataset (we will call this dataset "voted"). We will end up wanting one column for successes (number of people who actually voted), and one column for total number of people who were elligible to vote (total registered). 

First, we note the variables we definitely would like to keep in our final dataset to answer our questions of interest. From the registered dataset, we would like to keep total registered voters. From the voted dataset, we would like to keep `total voters`. We would also like to keep `sex_code`, `age`, and `county_desc` from both datasets. Additional variables of interest include `voted_party_cd`, `race_code`, `ethnic_code` which appear in both datasets, so we can merge on these. From the voted dataset, we will not keep `voting_method`, and `voting_method_desc` since they do not appear in the registered dataset. From the prompt, we also note that we use the `voted_party_cd` variable in the voted dataset, and not the `party_cd` variable.

We note that in the registered dataset, `precinct_abbrv` and `vtd_abbrv` will not be useful as they contain over 1000 factors each. `update_date` is also only NAs, so we will exclude that variable as well. We also will exclude `election_date` and `stats_type` as indicated in the prompt. 

```{r}
registered <- registered %>%
  dplyr::rename(total_registered = total_voters) %>%
  dplyr::rename(voted_party_cd = party_cd) %>%
  mutate(total_registered = as.numeric(total_registered))

voted <- voted %>%
  select(-voting_method_desc) %>%
  mutate(total_voters = as.numeric(total_voters))
```

Overall, in the voted dataset, we would like to keep the following variables: `county_desc`, `age`, `voted_party_cd`, `race_code`, `ethnic_code`, `sex_code`, `total_voters`, so we will aggregate to this level using a dplyr::group_by(). Overall, in the registered dataset, we would like to keep the following variables: `county_desc`, `voted_party_cd`, `race_code`, `ethnic_code`, `sex_code`, `age`, `total_registered`, so we will again aggregate to this level using a dplyr::group_by(). After this, we can merge the datasets.

```{r}
v <- voted %>%
  group_by(county_desc, age, voted_party_cd, race_code, ethnic_code, sex_code) %>%
  summarize(total_voters = sum(total_voters))

r <- registered %>%
  group_by(county_desc, voted_party_cd, race_code, ethnic_code, sex_code, age) %>%
  summarize(total_registered = sum(total_registered))

vote <- r %>%
  left_join(v, by = c("county_desc", "age", "voted_party_cd", "race_code",
                      "ethnic_code", "sex_code")) %>%
  mutate(total_registered = as.numeric(total_registered), 
         total_voters = as.numeric(total_voters))

```

We note that after merging, we have 6081 missing observations for `total_voters`.  This is because there were demographic/geographic groups in registered dataset that did not exist in the voted dataset. This likely means that these demographic groups who registered to vote, did not actually vote, since there are no values of `total_voters` = 0 in the voted dataset. Thus, we impute 0 for the missing values of `total_voters` to indicate that these demographic groups did not vote. 

Finally, we notice that some counties have instances where the total number of voters is larger than the total number of registered voters (12 observations in total), which is impossible and will not work in our model. For these counties, we reduce the total number of voters to be equal to the total number of registered voters. 

```{r, include=FALSE}
# apply(vote, 2, function(x) sum(is.na(x))) %>%
#   kbl()

vote <- vote %>%
  mutate(total_voters = as.numeric(total_voters), 
         total_registered = as.numeric(total_registered), 
         county_desc = as.factor(county_desc), 
         voted_party_cd = as.factor(voted_party_cd), 
         race_code = as.factor(race_code), 
         sex_code = as.factor(sex_code), 
         ethnic_code = as.factor(ethnic_code), 
         age = as.factor(age)
         ) %>%
  mutate(total_voters = case_when(
    total_voters > total_registered ~ total_registered, 
    TRUE ~ total_voters
  )) %>%
  mutate(total_voters = ifelse(is.na(total_voters), 0, total_voters))
```


```{r, include=F}
# test that all the NAs in the final dataset in voted_party_cd is correct - it is
registered %>%
  filter(county_desc == "ALAMANCE", voted_party_cd == "CST", race_code == "B", 
         ethnic_code == "NL", sex_code == "F", age == "Age 18 - 25")

voted %>%
  filter(county_desc == "ALAMANCE", voted_party_cd == "CST", race_code == "B", 
         ethnic_code == "NL", sex_code == "F", age == "Age 18 - 25")
```

### Final dataset: 51,906 observations (after imputing 0 for NA values in `total_voters`)

Finally, we select a random sample of 30 counties to continue our modeling with. The 30 counties we choose are: Hyde, Edgecombe, Haywood, Clay, Montgomery, Durham, Cabarrus, Cherokee, Duplin, Orange, Bladen, Sampson, Transylvania, Davie, Surry, Stanly, Watauga, Caldwell, Anson, Robeson, Beaufort, Pender, Graham, Stokes, Martin, Lenoir, Hertford, Wilson, Randolph, and New Hanover. 

The final dataset after sampling 30 counties has 14,935 observations after deleting missing values.

```{r}
set.seed(99)
county_samp <- as.character(sample(unique(vote$county_desc), size = 30))

# this is a random sample, save it so report is reproducible
county_samp <- c("HYDE", "EDGECOMBE", "HAYWOOD", "CLAY", "MONTGOMERY", "DURHAM", "CABARRUS", "CHEROKEE", "DUPLIN", "ORANGE", "BLADEN", "SAMPSON", "TRANSYLVANIA", "DAVIE", "SURRY", "STANLY", "WATAUGA", "CALDWELL", "ANSON", "ROBESON", "BEAUFORT", "PENDER", "GRAHAM", "STOKES", "MARTIN", "LENOIR", "HERTFORD", "WILSON", "RANDOLPH", "NEW HANOVER")

vote <- vote %>%
  filter(county_desc %in% c(county_samp))

```


# EDA

First we de-aggregate the dataset so that we can do some EDA on it, meaning we expand the dataset to have our outcome variable, whether someone voted or not (`voted`) be 0 or 1, rather than having an aggregated count. This increases our dataset to 1,590,202 observations. Note that we will use the aggregated dataset for modeling since having over 1 million observations is computationally infeasible.

```{r}
# create a 0 variable for people who didnt vote
vote <- vote %>%
  mutate(total_nonvoters = total_registered - total_voters)

# elongate vote dataset to make it not aggregate
vote_long <- vote %>%
  pivot_longer(cols = c("total_voters", "total_nonvoters"), names_to = "voted", values_to = "freq") %>%
  mutate(voted = ifelse(voted == "total_voters", "Yes", "No")) %>%
  mutate(obs = map(freq, ~rep_len(1, .x))) %>%
  unnest() %>%
  select(-freq, -obs)
```

## Examine outcome variable: `voted`

Our outcome variable, `voted`, is binary, and equals "Yes" if the person in the dataset voted, and "No" if they did not. We see that there are 1,190,593 people who voted and 402,165 people who did not vote.

```{r, eval=FALSE}
table(vote_long$voted)
```

## Research question 2: random intercept for county

Our 2nd research question asks us to examine whether the overall probability or odds of voting differed by county in 2020. This points to inclusion of a random intercept by county. We can see that there is a small amount variation in the proportion of registered voters who voted by county. We will include this variable as our grouping variable in our logistic hierarchical model based on our research questions. 

```{r,fig.height=4,fig.width=5}
ggplot(vote_long) + 
  geom_bar(aes(x = county_desc, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion", 
       title = "Proportion voting in each county") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        axis.title.x = element_blank())

```


## Research question 1: Examine fixed effects

Our first research questions asks to examine how different demographic subgroups voted, and how turnout for males compared to turnout for females after controlling for other predictors. This points to examination of different fixed effects, with a emphasis on examining gender. In our model selection process, we will test all fixed effects, but we use EDA to examine which fixed effects may be most related to our outcome. `age`, `voted_party_cd`, `race_code`, and `ethnic_code` seem to vary the most among their levels with the proportion of those who voted. Note that we will also definitely include `sex_code` as this is explicitly in our research question.  

```{r, fig.width = 8, fig.height=4}
names <- vote_long %>% 
  select(-county_desc, -total_registered) %>%
  colnames

names <- names[-1]

vote_long %>% select(-county_desc, -total_registered) %>%
  pivot_longer(-voted) %>% 
  mutate(name = factor(name, levels = names)) %>%
  mutate(voted = as.factor(voted)) %>%
  mutate(type = "barplot") %>%
  ggplot() + 
  geom_bar(
    aes(x = value, fill = voted),
    position = "fill",
    data = ~subset(.x, type == "barplot")
  ) +
  facet_wrap(~name, scale = "free", ncol = 3) +
  xlab("") +
  ylab("Proportion") +
  labs(title = "Proportion of registered voters who voted (Yes) \nand did not vote (No) by variable") +
  guides(fill=guide_legend(title="Value")) +
  theme(strip.text.x = element_text(size = 5), 
        legend.position = "right", 
        axis.text.x = element_text(angle = 90, size = 6, hjust=1))
```


## Research questions 3 and 4: interactions 

Research question 3 asks how the turnout rates different between males and females for different parties, and question 4 asks how turnout rates differ between age groups for the different parties. These point to `sex_code:voted_party_cd` and `age:voted_party_cd` interactions, respectively. We will thus include these interactions in our final model. According to the below plots, we can see there is some initial evidence of different turnout rates of genders and ages by parties, and we will explore this further in our model.  

```{r, fig.height = 4, message=F}
sex_code_int <- ggplot(vote_long) + 
  geom_bar(aes(x = sex_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~voted_party_cd, ncol= 6)

age_int <- ggplot(vote_long) + 
  geom_bar(aes(x = age, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5)) + 
  facet_wrap(~voted_party_cd, ncol = 6)

patchwork <- sex_code_int / age_int +
  plot_layout(guides = "collect", widths = c(1, 1))

patchwork + plot_annotation(
  title = "Proportion of registered voters who voted and did not vote \nvs. gender and age by party")
```

We also assess other potential interactions, and find that `sex_code:race_code` and  `voted_party_cd:race_code` could be helpful. 
With that being said, we decide to exclude all two-way interactions that do not 
address the state research questions. Additional two-way interactions increased the difficulty 
of fitting a model and greatly increase the difficulty of interpreting our results.


## Assessing random slopes 
We also assess potential random slopes of our variables by county. We do see some heterogeneity of voter turnout by `ethnic_code` among the counties  (See appendix for plots). Ultimately we decided to not test for random slopes. When only 
considering random intercepts, we encountered convergence issues with frequentest model 
and excessive runtimes with the Bayesian implementation. Therefore, the addition of 
random slopes would only increase the difficulty of fitting a model and did not specifically 
address our research questions. 


# Model selection and specification

We first tried to do exhaustive search selection using BIC, starting with the base model of the random intercept for county, and a fixed effects for `sex_code`, `age`, `voted_party_cd`, `sex_code:voted_party_cd`, and `age:voted_party_cd` based on our research questions. We aimed to use BIC to select other potential fixed effect predictors and test all possible 2-way interactions, however, none of our models converged. BIC is thus an unreliable measure to compare models, and we choose to use our model that incorporates elements from the research question as our final model. As mentioned we do not fit any random slopes due to convergence issues. Due to these convergence issues we ultimately use a Bayesian implementation.

```{r, eval=FALSE}
exhaustive_search <- function(raw_model, vars, data) {
  y_name <- deparse(raw_model[[2]], width.cutoff = 500L)
  group_name <- deparse(raw_model[[3]], width.cutoff = 500L)
  id <- 0 : (2^length(vars) - 1)
  construct_model <- function(.id){
    subset <- (.id %/% 2^(0:(length(vars) - 1))) %% 2 == 1
    if (all(subset == F)){
      RHS <- paste0(c("1", group_name), collapse = ' + ')
    }
    else {
      RHS <- paste0(c(vars[subset], group_name), collapse = ' + ')
    }
    paste(y_name, RHS, sep = ' ~ ')
  }
  run_model <- function(.id){
    model_str <- construct_model(.id)
    model_formula <- as.formula(model_str)
    res <- glmer(model_formula, data = data, family = binomial("logit"))
    return (summary(res)$AICtab)
  }
  models <- sapply(id, construct_model)
  result <- pbapply::pbsapply(id, run_model)
  bind_cols(
    model = models,
    as.data.frame(t(result))
  )
}
```

```{r, message = F, echo = F, cache = T, eval=FALSE}
# may take 2-3 hrs. result saved to 'ex_result_int.RDS'
ex_result_int <- exhaustive_search(
  raw_model <- cbind(total_voters, total_nonvoters) ~ sex_code + age + voted_party_cd + 
    sex_code:voted_party_cd + age:voted_party_cd + (1 | county_desc),
  vars = c("race_code", "ethnic_code", "sex_code:age", "voted_party_cd:race_code"),
  data = vote
)
saveRDS(ex_result_int, "ex_result_int.RDS")
```

```{r, echo = F, eval=F}
ex_result_int <- readRDS("case_study_02/ex_result_int.RDS")
ex_result_int %>% arrange(BIC) %>% head(10) %>%
  select(1, 3) %>%
  kbl(caption = "Exhaustive search of fixed effects using BIC") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                latex_options = "HOLD_position")
```

Our final model is thus: 

$$
\begin{aligned}
& y_{ij} \mid N_{ij}, \pi_{ij} \stackrel{\mathrm{iid}}{\sim} \mathrm{Bin}(N_{ij}, \pi_{ij}),
\quad
\mathrm{logit}(
\pi_{ij}
) = \boldsymbol{x}_{ij}^\mathrm{T}\boldsymbol{\beta} + b_i,
\quad
b_i \sim N(0, \sigma^2) \\
& \beta \sim N(\boldsymbol{\mu}, \Sigma), \quad \sigma^2 \sim IG(\nu_0/2, \nu_0\sigma_0^2/2)
\end{aligned}
$$

where $y_{ij}$ is the number of people voted, and $N_{ij}$ is the number of total voters for group $j$ in county $i$ (random intercept indexed by $i$). Each $\boldsymbol{x}_{ij}$ consists of 86 elements - the first one being 1 for intercept, and the next 85 elements being the value of the covariates (in one hot form) for group $j$ in county $i$, namely each possible value of `race_code`, `ethnic_code`, `sex_code`, `age`, `voted_party_cd`, `sex_code:age`, `voted_party_cd:race_code`, `voted_party_cd:sex_code`, `voted_party_cd:age` except baseline. It should be noted 
that $\Sigma$ is a diagonal matrix as `brms` only allows indpendent priors on the fixed effect coefficients.



# Results and Interpretation 

```{r}
bayesian_base_model <- readRDS("bayesian_result.RDS")
```

# Interpretations 

## Main effects
Our estimated intercept is -0.04. This is the expected log odds of someone voting, across all counties in our sample,
who is Asian, ages 26-40, female, and did not specify their ethnicity. That means
the odds of voting is approximately 1. The estimate, clearly is not significant and contains 1 
in the posterior credible interval (on odds scale).

For the purposes of interpretation, statements of significance imply that the 95% posterior 
credible interval did not include 0. The following highlight significant results 
concerning the relationship between voting odds and various demographics.

* On average, we expect the odds of voting for someone between the ages 18-26 to decrease
by 56% when compared to some that is Ages 26-40.  

* On average, we expect the odds of voting for someone who is White to increase
by 46% when compared to some who is Asian. We identified a similar relationship for 
someone registered with an undesignated race.   

* On average, we expect the odds of voting for males to increase
by 42% when compared to females.

* On the other hand, we expect the odds of voting for someone who is registered under two or more races to decrease
by 8% when compared to some who is Asian. We identified a similar relationship for 
someone registered as black.  

* On average, we expect the odds of voting for someone who is not Hispanic or Latino to increase
by 14% when compared to some who has an undesignated ethnicity. Someone who is Latino or 
Hispanic sees an expected decrease in odds of voting of 25% when compared to someone 
with an undesignated ethnicity.  



```{r}
set.seed(5562)
new_data_sex <- expand.grid(unique(vote$sex_code),
                            unique(vote$voted_party_cd)) %>% 
  rename(sex_code = Var1, voted_party_cd = Var2) %>% 
  mutate(race_code = as.factor(sample(unique(vote$race_code),1)),
         ethnic_code = as.factor(sample(unique(vote$ethnic_code),1)),
         age = as.factor(sample(unique(vote$age),1)),
         county_desc = as.factor(sample(unique(vote$county_desc),1)),
         sex_code = as.factor(sex_code),
         total_registered = 1
         )
new_data_sex$race_code <- droplevels(new_data_sex$race_code)

new_data_sex$probs <- predict(bayesian_base_model, new_data_sex, type="response")[,1]

p1 <- ggplot(new_data_sex, aes(x = sex_code, y = probs, fill = sex_code)) +
  geom_bar(stat = "Identity") + facet_wrap(~voted_party_cd) + 
  labs(title = "Effect of party affiliation on\nsex and probability of voting",
       y = "Probability of Voting", x = "") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + 
  guides(fill=guide_legend(title="Sex"))
```

```{r}
set.seed(5562)
new_data_age <- expand.grid(unique(vote$age),
                            unique(vote$voted_party_cd)) %>% 
  rename(age = Var1, voted_party_cd = Var2) %>% 
  mutate(race_code = sample(unique(vote$race_code),1),
         ethnic_code = sample(unique(vote$ethnic_code),1),
         sex_code = sample(unique(vote$sex_code),1),
         county_desc = sample(unique(vote$county_desc),1),
         total_registered = 1
         )
new_data_age$race_code <- droplevels(new_data_age$race_code)

new_data_age$probs <- predict(bayesian_base_model, new_data_age, type="response")[,1]

p2 <- ggplot(new_data_age, aes(x = age, y = probs, fill = age)) +
  geom_bar(stat = "Identity") + facet_wrap(~voted_party_cd) + 
  labs(title = "Effect of party affiliation on\nage group and probability of voting",
       y = "", x = "") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + 
  guides(fill=guide_legend(title="Age Group"))
```

## Interaction Effects
```{r, fig.height=4, fig.width=8}
p1 + p2
```

From the above plots we see that older age groups consistently have a higher probability 
of voting across all parties. There is some variation across parties in probability of voting between 
age groups 26-40 and 18-25. Overall, there doesn't appear to be a strong interaction effect
between age group and voter party. We see a similar pattern for gender and voter party. 
Undesignated gender consistently had the highest probability of voting, while male and females
often had a similar probability of voting across parties with the exception of CST and 
DEM.



```{r, fig.height=4, fig.width=5}
# dotplot
library(lattice)
tmp <- bayesian_base_model %>%
  spread_draws(r_county_desc[county_desc,]) %>%
  median_qi(`Group Means` = r_county_desc)
sorted_county_desc <- tmp %>%
  arrange(`Group Means`, desc = T) %>% pull(county_desc)
p1 <- tmp %>%
  ggplot(aes(y = factor(county_desc, levels = sorted_county_desc), 
             x = `Group Means`, xmin = .lower, xmax = .upper)) +
  geom_pointinterval(orientation = "horizontal", fatten_point = .8) + 
  labs(title = "Random effect of log odds by County", 
       y = "County", 
       x = "Log Odds") + 
    geom_vline(aes(xintercept = 0)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8)) +
  coord_flip()
p1
```

The interval plot demonstrates a large amount of heterogeneity in the log odds 
of voting across our sampled counties. For a registered voter who is Asian, 
between ages 26-40, female, and did not specify their ethnicity, the odds of voting 
decreases by more than 32.9% if they are in Robeson county and increases by more 
than 22% if they are in Cabarrus county.

# ROC Curve for Model Evaluation
We plotted an ROC curve to evaluate our model and see how well it does at predicting
the whether or not a particular voter voted. From the curve below, you can see that our
model does better than simply taking random guesses for whether or not a voter belonging
to a particular demographic within a particularly voted or not. Furthermore, the AUC from
our ROC curve is 0.681, which indicates that our model does a decent job of prediction since
the value is greater than 0.5, which is what the AUC would be had we simply taken random guesses
as to whether an individual voted or not. Thus, our model seems to be useful in predicting whether
or not individuals of certain demographics in various counties in NC voted or not.
```{r, fig.height=4, fig.width=5}
set.seed(4)
test_subset <- filter(vote_long, county_desc %in% county_samp)
test_subset <- test_subset[sample(c(1:nrow(test_subset)), 3000), ]
test_subset <- test_subset %>% 
  select(sex_code, age, voted_party_cd, race_code, ethnic_code, county_desc, voted) %>%
  mutate(race_code = as.factor(race_code),
         age = as.factor(age),
         sex_code = as.factor(sex_code),
         ethnic_code = as.factor(ethnic_code),
         voted_party_cd = as.factor(voted_party_cd),
         total_registered = 1,
         voted = if_else(voted == "Yes", 1, 0),
         predicted = NA)
test_subset$race_code <- droplevels(test_subset$race_code)
set.seed(4)
test_subset[["predicted"]] <- predict(bayesian_base_model, newdata = test_subset, type="response")[,1]
roc_data <- roc(data = test_subset, response = "voted", predictor = "predicted")
plot.roc(roc_data)
auc(roc_data)
```

# Appendix

# Code Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```

# Plot Appendix

Testing Interactions 

```{r}
# Interactions
####### sex:code ###############################################################
# sex_code:race_code
ggplot(vote_long) + 
  geom_bar(aes(x = sex_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~race_code, ncol= 4)


# sex_code:ethnic_code
ggplot(vote_long) + 
  geom_bar(aes(x = sex_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~ethnic_code, ncol= 4)

# sex_code:age
ggplot(vote_long) + 
  geom_bar(aes(x = sex_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~age, ncol= 4)

######## party_cd ##############################################################
# party_cd:race_code
ggplot(vote_long) + 
  geom_bar(aes(x = voted_party_cd, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~race_code, ncol= 4)

# party_cd:ethnic_code
ggplot(vote_long) + 
  geom_bar(aes(x = voted_party_cd, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~ethnic_code, ncol= 4)

###### ethnic_code #############################################################
# ethnic_code:race_code
ggplot(vote_long) + 
  geom_bar(aes(x = ethnic_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~race_code, ncol= 4)

# ethnic_code:age
ggplot(vote_long) + 
  geom_bar(aes(x = ethnic_code, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5), 
        legend.position = "none") + 
  facet_wrap(~age, ncol= 4)
```

Random slopes


```{r}
# MOVE TO APPENDIX
# Random slopes 

# voted_party_cd
ggplot(vote_long) + 
  geom_bar(aes(x = county_desc, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5)) + 
  facet_wrap(~voted_party_cd, ncol = 4)

# race_code
ggplot(vote_long) + 
  geom_bar(aes(x = county_desc, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5)) + 
  facet_wrap(~race_code, ncol = 4)

# ethnic_code
ggplot(vote_long) + 
  geom_bar(aes(x = county_desc, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5)) + 
  facet_wrap(~ethnic_code, ncol = 4)

# sex_code
ggplot(vote_long) + 
  geom_bar(aes(x = county_desc, fill = voted), 
           position = "fill") + 
  labs(y = "Proportion") +
  theme(axis.text.x = element_text(angle = 90, vjust = -0.5, hjust=1), 
        plot.title = element_text(size = 10, hjust = 0.5)) + 
  facet_wrap(~sex_code, ncol = 4)
```



Table of Main Effects
```{r, eval=F}
base_model <-glmer(
              cbind(total_voters, total_nonvoters) ~ sex_code + age + voted_party_cd +  
              sex_code:voted_party_cd + race_code + ethnic_code +
              age:voted_party_cd + (1 |county_desc), data = vote,
              family = binomial("logit"), 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
              )

saveRDS(base_model, "base_model.RDS")
```

```{r}
base_model <- readRDS('base_model.RDS')
coef(summary(base_model)) %>%
  knitr::kable(caption = "Fixed effect estimates on log odds scale") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                latex_options = "HOLD_position")
```

## Bayesian Modeling
```{r, eval=FALSE}
bayesian_base_model <-brm(
              total_voters | trials(total_registered) ~ sex_code + age + voted_party_cd +  
              sex_code:voted_party_cd + race_code + ethnic_code +
              age:voted_party_cd + (1 |county_desc), data = vote,
              family = binomial("logit"),
              chains = 1)

```

Table of Main Effects
```{r}
summary_pars <- summary(bayesian_base_model)$fixed[, c(1,3,4)]
# rownames(summary_pars) <- c("Intercept (grand mean)", "numbeds", "sigma", "sd_control", "sd_state")
colnames(summary_pars) <- c("Est", "Lwr", "Upr")
summary_pars %>%
  knitr::kable(caption = "Estimated posterior parameters", 
               digits = 3)
```


```{r}
bayesian_<- readRDS("bayesian_result.RDS")
```

