---
title: "Lab 9"
author: "Marc Brooks, Bo Liu, Shirley Mathur, Aasha Reddy"
date: "10/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(lme4)
```


```{r}
# load data
badgers <- read.csv("data/BadgersFarmSurveysNoNA.csv", sep = "")
```

# Introduction and Data

Management of badgers on Welsh and British farms is a controversial topic. While some argue that culling badger populations is necessary to control the spread of bovine tuberculosis, others argue that badgers are not the major cause of the spread of disease and that culling is inhumane, especially when badger vaccination strategies are an option.

We consider data from observational surveys of 36 farms over a period of three years. The data are available on Sakai under Resources, in the BadgersFarmSurveysNoNA.txt file. Each farm (identified by the variable farm_code_numeric) was observed up to eight times (once per season for a two year period). The primary outcome, Signs_in_yard, is whether badger activity (e.g., carcasses of their prey, badger feces, indications of digging) was present in the farmyard.

## Research questions
* What factors relate to presence of badger activity in the farmyard?
* Estimate the correlation over time in badger activity and farm-specific heterogeineity in the tendency to have badger activity.

# EDA



