---
title: "Lab 9"
author: "Marc Brooks, Bo Liu, Shirley Mathur, Aasha Reddy"
date: "10/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# load libraries
library(tidyverse)
library(lme4)
library(patchwork)
library(lubridate)
library(brms)
library(rstan)
library(knitr)
library(kableExtra)
library(corrplot)
library(cAIC4)
```


```{r}
# load data
badgers <- read.csv("../data/BadgersFarmSurveysNoNA.csv", sep = "")
```


# Introduction and Data

Management of badgers on Welsh and British farms is a controversial topic. While some argue that culling badger populations is necessary to control the spread of bovine tuberculosis, others argue that badgers are not the major cause of the spread of disease and that culling is inhumane, especially when badger vaccination strategies are an option.

We consider data from observational surveys of 36 farms over a period of three years. The data are available on Sakai under Resources, in the BadgersFarmSurveysNoNA.txt file. Each farm (identified by the variable farm_code_numeric) was observed up to eight times (once per season for a two year period). The primary outcome, Signs_in_yard, is whether badger activity (e.g., carcasses of their prey, badger feces, indications of digging) was present in the farmyard.

## Research questions

1) What factors relate to presence of badger activity in the farmyard?
2) farm-specific heterogeneity in the tendency to have badger activity
3) Estimate the correlation over time in badger activity

# EDA

## Examine outcome variable: signs_in_yard

First we examine how many 0's and 1's are in the outcome variable, signs_in_yard. We see that there are 230 0's and 43 1's. 


## Research Question 1: Examine fixed effects

Our first research question asks us to examine factors related to presence of badger activity in the farmyard. Thus, we use EDA to exmaine which fixed effects to include in our model. We first note that we will not consider Season, time and survey, as each sample for each farm was taken at a different time point. Thus, the farm variable already accounts for a time effect. We would be unable to include all of these variables in our model since there would be multicollinearity. 

We examine plots of other predictors vs. our outcome (Signs_in_yard) to decide which variables we will consider in our model selection process (exhaustive search using BIC). We find the following variables show a difference between the counts of 0s and 1s in our outcome, and we consider them in our exhaustive search (see plots in appendix): no_setts_in_field, no_active_setts_in_field, no_cattle_cattle_in_buildings, accessible_feed__store_present, Hay straw, concentrates, Sugar beet, Molasses. 

We find that no_active_setts_in_fields is the most important variable with respect to our outcome, Signs_in_yard. 

```{r, fig.align='center'}
# No setts in field
setts <- ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_active_setts_in_fields, fill = Signs_in_yard)) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard")

# No active setts in fields
act_setts <- ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_active_setts_in_fields, fill = Signs_in_yard)) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard")

patchwork <- setts + act_setts

patchwork + plot_annotation(
  title = "Signs_in_yard vs. no. settlements and active settlements")

```


```{r}
# No_cattle_in_buidlings_yard
cattle_build <- ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_cattle_in_buidlings_yard, fill = Signs_in_yard)) +
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No_cattle_in_buidlings_yard")

afsp <- ggplot(badgers, aes(x = Accessible_feed_store_present, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Accessible_feed_store_present vs. Signs_in_yard") + 
  theme(legend.position="bottom")

haystraw <- ggplot(badgers, aes(x = HayStraw, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "HayStraw vs. Signs_in_yard") + 
  theme(legend.position="bottom")

conc <- ggplot(badgers, aes(x = Concentrates, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Concentrates vs. Signs_in_yard") + 
  theme(legend.position="bottom")

sugarbeet <- ggplot(badgers, aes(x = Sugarbeet, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Sugarbeet vs. Signs_in_yard") + 
  theme(legend.position="bottom")

molasses <- ggplot(badgers, aes(x = Molasses, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Molasses vs. Signs_in_yard") + 
  theme(legend.position="bottom")

patchwork <- (cattle_build + afsp + haystraw) / (conc + sugarbeet + molasses)

patchwork + plot_annotation(
  title = "Signs_in_yard vs. other fixed effects")

```

## Research Question 2 and 3: Random intercept for farm

Our 2nd research question asks us to examine farm-specific heterogeneity in the tendency to have badger activity. This points to including a random intercept for farm in our model. This also accounts for correlation over time in badger activity (research question 3), as each farm's measurements were taken at a different point in time. Since our ICC is relatively high, we find that badger activity is relatively correlated over time. 


# Model selection and specification

Next, We do step-wise selection using BIC, starting with the base model of the random intercept for farm, and a fixed effect for no_active_setts_in_fields as this variable appeared to be most important from EDA. Based on step-wise selection, we choose our final model to be the following (see test output in Appendix):


$$
logit(E(Y_{ij}|b_{i})) = X'_{ij}\beta + b_i
$$
$$
b_i \sim N(0, \sigma^2)
$$
Where $Y_{ij}$ is the Signs_in_yard indicator for observation $j$ in farm $i$ (random intercept indexed by $i$). In $X'$, we have 2 columns, one for the intercept and one for our fixed effect No_setts_in_field. 


# Interpretation and Results 

### Research Question 1: Fixed Effects

Our first research question asks us to examine factors related to the presence of badger activity in the farmyard. Through step-wise selection using BIC, we find that the model with only No_setts_in_fields led to the lowest BIC. The below table shows the estimates of the fixed effects on a logit scale, which we can interpret on the odds by exponentiating as follows: 

* For a 1 unit increase in the number of active settlements in the field, we expect the odds of the presence of signs of badgers in the farmyard to increase by 45% (exp(0.37) = 1.45).
* Intercept: In the absence of any settlements in the field, we expect the odds of the presence of signs of badgers in the farmyard to decrease by 99% (exp(-5.03) = 0.007).

```{r}
m1 <- glmer(Signs_in_yard ~ No_setts_in_fields + 
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
```

```{r}
coef(summary(m1)) %>%
  knitr::kable(caption = "Fixed effect estimates on logit scale", 
               digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                latex_options = "HOLD_position")
```

### Research Question 2: Random Intercept

```{r, fig.height=5}
# dotplot
library(lattice)
dotplot(ranef(m1, condVar = TRUE), font.size = 5, rotate = TRUE, 
        main = FALSE)$farm_code_numeric
```

Our research question 2 asks us to examine farm-specific heterogeneity, so we can examine our random intercept for farm. From the dotplot, we see that there are some differences across the farms. However, even though there are differences, the confidence intervals are pretty wide. This could be due to small sample sizes, though some confidence intervals do not include 0. 

```{r}
as.data.frame(VarCorr(m1)) %>%
  select(1, 4) %>%
  rename(var = vcov) %>%
  knitr::kable(caption = "Variance estimates") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                latex_options = "HOLD_position")
```

### Research Question 3: ICC

The 3rd research question asks us to estimate the correlation over time in badger activity, and we thus examine the ICC. We find that our ICC is ~0.6, which measures the correlation of observations within each farm. This also accounts for correlation over time in badger activity, as each farm's measurements were taken at a different point in time. As such, this study is a kind of longitudinal study. Since our ICC is relatively high, we find that badger activity is relatively correlated over time. 

$$
ICC = \frac{\sigma^2}{\sigma^2 + \frac{\pi^2}{3}} = 0.598
$$


# Appendix 

# Code Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```


# Plot Appendix

## EDA 

```{r}
# No setts in field
ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_setts_in_fields, fill = Signs_in_yard)) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No_setts_in_fields")
```


```{r}
# No active setts in fields
ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_active_setts_in_fields, fill = Signs_in_yard)) + 
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No__active_setts_in_fields")
```

```{r}
# No_buildings
ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_buildings, fill = Signs_in_yard)) +
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No_buildings")
```


```{r}
# No_cattle_in_buidlings_yard
ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_cattle_in_buidlings_yard, fill = Signs_in_yard)) +
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No_cattle_in_buidlings_yard")
```

```{r}
# No_cattle_in_buidlings_yard
ggplot(badgers, aes(x = as.factor(Signs_in_yard), y = No_cattle_in_buidlings_yard, fill = Signs_in_yard)) +
  geom_boxplot() + 
  theme(legend.position = "none") + 
  labs(x = "Signs in Yard", 
       title = "Signs_in_yard vs. No_cattle_in_buidlings_yard")
```

```{r}
# season
table(badgers$Signs_in_yard, badgers$Season)
```


```{r}
# binary predictors
prop.table(table(badgers$Signs_in_yard, badgers$Accessible_feed_present))
prop.table(table(badgers$Signs_in_yard, badgers$Accessible_cattle_house_present))
prop.table(table(badgers$Signs_in_yard, badgers$Accessible_feed_store_present))
prop.table(table(badgers$Signs_in_yard, badgers$Grass_silage))
prop.table(table(badgers$Signs_in_yard, badgers$Cereal_silage))
prop.table(table(badgers$Signs_in_yard, badgers$HayStraw))
prop.table(table(badgers$Signs_in_yard, badgers$Cereal_grains))
prop.table(table(badgers$Signs_in_yard, badgers$Concentrates))
prop.table(table(badgers$Signs_in_yard, badgers$Proteinblocks))
prop.table(table(badgers$Signs_in_yard, badgers$Sugarbeet))
prop.table(table(badgers$Signs_in_yard, badgers$Vegetables))
prop.table(table(badgers$Signs_in_yard, badgers$Molasses))
```


```{r}
ggplot(badgers, aes(x = Accessible_feed_present, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Accessible_feed_present vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Accessible_cattle_house_present, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Accessible_cattle_house_present vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```

 
```{r}
ggplot(badgers, aes(x = Accessible_feed_store_present, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Accessible_feed_store_present vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```

```{r}
ggplot(badgers, aes(x = Grass_silage, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Grass_silage vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```

```{r}
ggplot(badgers, aes(x = Cereal_silage, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Cereal_silage vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = HayStraw, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "HayStraw vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```

```{r}
ggplot(badgers, aes(x = Cereal_grains, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Cereal_grains vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Concentrates, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Concentrates vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Proteinblocks, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Proteinblocks vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Sugarbeet, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Sugarbeet vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Vegetables, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Vegetables vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```


```{r}
ggplot(badgers, aes(x = Molasses, fill = as.factor(Signs_in_yard))) + 
  geom_bar(position = "fill") + 
  labs(title = "Molasses vs. Signs_in_yard") + 
  theme(legend.position="bottom")
```

## Model selection (step-wise BIC)



```{r}
# APPENDIX

# BIC 169.5    
m1 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + 
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m1)

# BIC 169.5  
m2 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_active_setts_in_fields +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m2)

# BIC 171.2
m3 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m3)

# BIC 173.0    
m4 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + Accessible_feed_store_present +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m4)

# BIC 172.3
m5 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + HayStraw +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m5)

# BIC 173.1
m6 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + Concentrates +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m6)

# BIC 172.4
m7 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + Sugarbeet +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m7)

# BIC 173.1 
m8 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + Molasses +
              (1 | farm_code_numeric), data = badgers, 
      family = binomial(link = "logit"), 
      control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m8)
```



